---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
```

# ggcirclepack and beyond - highly experimental

<!-- badges: start -->
<!-- badges: end -->

Critical moment in my extension journey started with TLP "extend your ability to extend".  

Message 1: you can be an extender 

--

Message 2: "where" aesthetics need not only be x and y.

--

aes(x0 = ?, y0 = ?)

--

And what ggplot2 actually looks at is computed for you (x and y)

---


Today I'll present on ggcirclepack. 

---

Background: ggplot2 grammar guide walk through (reference from data to viz gallery)

```{r}
knitr::include_url("https://evamaerey.github.io/ggplot2_grammar_guide/geoms_single_series.html#77")

```


--

This didn't feel like powerful ggplot2 experience


--

What would ggcirclepack look like? 

---


Here, the id of the circle isn't defined by the center, x0y0, but by its id and area (an algorithm used under the hood to find x0y0 and x and y for perimeter)

--

Working with compute_panel(), ggols



---


## Installation

You can install the development version of ggcirclepack from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("EvaMaeRey/ggcirclepack")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ggcirclepack)
library(ggplot2)
library(magrittr)
library(dplyr)

gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  geom_polygon_circlepack(alpha = .5) + 
  coord_equal()

last_plot() +
  aes(fill = continent) + 
  labs(title = "gapminder 2002 countries")

last_plot() +
  aes(area = pop) + 
  geom_text_circlepack()

last_plot() +
  facet_wrap(facets = vars(continent))

last_plot() + 
  scale_size_continuous(range = c(0, 4)) + 
  theme(legend.position = "none")

last_plot() + 
  aes(area = gdpPercap*pop)

last_plot() + 
  aes(area = gdpPercap)



```


---

Wish list for ggcirclepack:

More computation under the hood for a count data case.  

```{r}
tidytitanic::tidy_titanic %>% 
  head()
```



```{}
tidytitanic::tidy_titanic() + 
  ggplot() + 
  aes(x = "all") + 
  geom_polygon_circlepack(alpha = .5) + 
  geom_text_circlepack(aes(label = afterstat(count))) + # automatically labels with count
  aes(linetype = sex) + 
  aes(color = age) + 
  aes(alpha = survived) + 
  facet_wrap(~class)
```

---

Other wish list item.. Include other circlepacking algorithms.

--

get feedback from pack circles author!  

---

### Extremely highly experimental: maps and atlasses.

--

aes(where = ?)
--
aes(country = ?)

aes(state = ?)

aes(fips = ?) # us county codes

aes(brain_segment = ?)

aes(tissue = ?)


---

# U.S. States 

- inheriting from geom_polygon
- 'computing' state xy perimeter based aes = my_state_var
- Ryan Miller tutorial start point.


```{r}
knitr::include_url("https://evamaerey.github.io/mytidytuesday/2022-11-04-brain/brain.html")
```

---


```{r}
knitr::include_url("https://evamaerey.github.io/mytidytuesday/2023-03-06-us-states/us_states.html")

```


---

# U.S. counties

- inheriting from geom_sf()
- 'computing' fips sf geometries column (xy perimeters) based on aes = my_fips_code
- Wilke example start point

```{r}
knitr::include_url("https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips_w_sf.html")
```




---

# Brain seg

- inheriting from geom_sf() 
- 'computing' tissue space based on aes = my_seg_name variable
- atlases from ggseg

```{r}
knitr::include_url("https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_sf.html")
```

