---
output: 
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", 
  warning = F
)
```

# ggcirclepack

<!-- badges: start -->
<!-- badges: end -->

circle pack is an experimental package that uses the {packcircles} package to handle circle packing computation. 

### Note to the reader

Your feedback is on this work is greatly appreciated.

Beyond the descriptions of our work, we interject comments on our hesitations 🤔 and areas that need some work 🚧, for your consideration marked with emoji.

Your help and feedback would be greatly appreciated on any of the questions...

- Are functions named intuitively? *'According to IBM studies, intuitive variable naming contributes more to code readability than comments, or for that matter, any other factor' McConnell, S. Code complete*
- Do functions work as you expect? 
- Is there rewriting that could make the code more concise?
- What tests should be performed?


# status quo *without* {ggcirclepack}: precomputation required to create two more data frames

```{r}
library(tidyverse)
gapminder::gapminder %>%  
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  select(country, pop) %>% 
  mutate(id = row_number()) ->  
df_w_id

packcircles::circleProgressiveLayout(df_w_id$pop,  
                                         sizetype = 'area') ->  
x0y0radius  

x0y0radius %>%  
  packcircles::circleLayoutVertices(npoints = 50) ->  
circle_outlines  

circle_outlines %>% 
  left_join(df_w_id) %>% 
  ggplot() +  
  aes(x = x, y = y) +  
  geom_polygon(colour = "black", alpha = 0.6) +  
  aes(group = id) +  
  aes(fill = pop) +  
  geom_text(data = cbind(df_w_id, x0y0radius),  
            aes(x, y, size = pop, label = country,  
                group = NULL, fill = NULL)) +  
  theme(legend.position = "none") +  
  coord_equal()
```


# Proposed UI

```{r example, out.width="33%", fig.show="hold", message=F, eval = F}
library(tidyverse)
library(ggcirclepack)

gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country, area = pop) +
  geom_circlepack() +               # draws packed circles 
  geom_circlepack_text() +          # labels at the center
  coord_fixed(ratio = 1)
```




# Package functions

## geom_circlepack_text (center)

### Step 1. compute panel

```{r compute_panel_circlepack_center}
compute_panel_aggregation <- function(data, scales, fun = sum){
  
    if(is.null(data$slice)){data$slice <- TRUE}

  data %>% 
    dplyr::filter(.data$slice) ->
  data
  
  grp_cols <-  c("id", "fill", "alpha", 
                 "colour", "group", "linewidth", 
                 "label", "size",
                 "linetype", "render")
  
  # Thanks June! https://github.com/teunbrand/ggplot-extension-club/discussions/15
  data %>% 
    group_by(group_by(pick(any_of(grp_cols)))) ->   
  data
  
  if(is.null(data$area)){data <- mutate(data, area = 1)}
  if(is.null(data$wt)){data$wt <- 1}
  
  data %>% 
    summarize(area = fun(.data$area*.data$wt), .groups = 'drop') ->
  data
    
  if(is.null(data$within)){data$within <- 1}

  data %>%   
    group_by(.data$within) %>% 
    mutate(prop = .data$area/sum(.data$area)) %>%
    mutate(percent = round(.data$prop*100)) -> 
  data
  
  data
  
}

#' compute_panel_circlepack_center
#'
#' @return
#' @export
#'
#' @examples
compute_panel_circlepack_center <- function(data, scales, fun = sum){

  # get aes names as they appear in the data
  
 data <- data |> compute_panel_aggregation(fun = fun) 
  
  
  data %>%   
    arrange(id)  -> # this doesn't feel very principled; motivation is when you go from no fill to color, preserves circle position...
  data  
  
  data %>%
    pull(area) %>%
    packcircles::circleProgressiveLayout(
      sizetype = 'area') %>%
    cbind(data) ->
  data
  
  
  if(!is.null(data$render)){
    
    data %>% 
      filter(.data$render) ->
    data
    
  }
  
  data
}

```

### Step 1.1 test compute

```{r}
gapminder::gapminder %>%
filter(continent == "Americas") %>%
  filter(year == 2002) %>%
  # input must have required aesthetic inputs as columns
  select(area = pop, id = country) %>%
  compute_panel_circlepack_center() %>%
  head()

gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  select(id = continent) %>% 
  compute_panel_circlepack_center()

gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  mutate( render = country == "Argentina") %>% 
  select(id = continent, render) %>% 
  compute_panel_circlepack_center()

gapminder::gapminder %>% 
  # filter(year == 2002) %>% 
  mutate( render = country == "Argentina") %>% 
  select(id = continent, render, area = pop) %>% 
  compute_panel_circlepack_center(fun = mean)



```

### Step 2 and 3 ggproto and geom

```{r geom_circlepack_text}
StatCirclepackcenter <- ggplot2::ggproto(`_class` = "StatCirclepackcenter",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("id"),
                                  compute_panel = compute_panel_circlepack_center,
                                  default_aes = ggplot2::aes(group = after_stat(id),
                                                             size = after_stat(area),
                                                             label = after_stat(id))
                                  )


#' Title
#'
#' @param mapping
#' @param data
#' @param position
#' @param na.rm
#' @param show.legend
#' @param inherit.aes
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
stat_circlepack_center <- function(mapping = NULL, data = NULL,
                                 geom = "text",
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    geom = geom, # inherit other behavior
    stat = StatCirclepackcenter, # proto object from Step 2
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}


#' Title
#'
#' @param mapping
#' @param data
#' @param position
#' @param na.rm
#' @param show.legend
#' @param inherit.aes
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
geom_circlepack_text <- stat_circlepack_center

```


### Step 4. test geom

```{r}
gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country, area = pop) +
  geom_circlepack_text(alpha = .5) + 
  coord_equal() + 
  labs(title = "gapminder 2002 countries")

last_plot() + 
  aes(render = pop > 20000000)


layer_data()
```





## geom_circlepack

### Step 1. compute_panel

```{r compute_panel_circlepack}
# Step 1
#' compute_panel_circlepack
#'
#' @param data
#' @param scales
#'
#' @return
#' @export
#'
#' @examples
compute_panel_circlepack <- function(data, scales, npoints = 50, fun = sum){

  data <- data |> compute_panel_aggregation(scales, fun)

  data %>%   
    arrange(id)  -> # this doesn't feel very principled; motivation is when you go from no fill to color, preserves circle position...
  data
  
  data$id = 1:nrow(data)

  data %>%
    pull(area) %>%
    packcircles::circleProgressiveLayout(
      sizetype = 'area') %>%
    packcircles::circleLayoutVertices(npoints = npoints) %>%
    left_join(data, by = join_by(id)) 

}
```

### Step 1.1. test compute

```{r}
gapminder::gapminder %>%
filter(continent == "Americas") %>%
  filter(year == 2002) %>%
  # input must have required aesthetic inputs as columns
  rename(id = country, area = pop) %>%
  compute_panel_circlepack() %>%
  head()
```

### Step 2 & 3 ggproto and geom

```{r geom_circlepack}
StatCirclepack <- ggplot2::ggproto(`_class` = "StatCirclepack",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("id"),
                                  compute_panel = compute_panel_circlepack,
                                  default_aes = ggplot2::aes(group = after_stat(id))
                                  )

#' Title
#'
#' @param mapping
#' @param data
#' @param position
#' @param na.rm
#' @param show.legend
#' @param inherit.aes
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
geom_circlepack <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCirclepack, # proto object from Step 2
    geom = ggplot2::GeomPolygon, # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

### Step 4. test geom

```{r}


gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  geom_circlepack(alpha = .5) + 
  geom_circlepack_text() +
  coord_equal() + 
  labs(title = "gapminder 2002 countries")

last_plot() +
  aes(fill = continent) + 
  labs(title = "from 5 continents")

last_plot() +
  aes(area = pop) + 
  labs(title = "with very different populations")

last_plot() +
  facet_wrap(facets = vars(continent)) + 
  labs(title = "faceting")

last_plot() + 
  scale_size_continuous(range = c(0, 4)) + 
  theme(legend.position = "none") + 
  labs(title = "remove legends")

last_plot() + 
  aes(area = gdpPercap*pop) + 
  labs(title = "and very different GDPs")

last_plot() + 
  aes(area = gdpPercap) + 
  labs(title = "and per capita GDPs 2002")

last_plot() + 
  aes(slice = continent == "Europe") + 
  facet_null() + 
  aes(fill = after_stat(area)) + 
  aes(label = paste0(country, "\n$", round(gdpPercap)))


gapminder::gapminder %>%
filter(year == 2002, 
       continent == "Europe") %>%
  ggplot() +
  aes(id = country, area = pop*gdpPercap) +
  geom_circlepack() + 
  geom_circlepack_text(vjust = 0, color = "grey75",
                       lineheight = .7) +
  aes(label = str_wrap(country, 10)) +
  coord_equal() + 
  scale_size(range = c(.5,4.5)) +
  aes(slice = continent == "Europe") +
  labs(title = "Percent of European Economy 2002") + 
  aes(fill = after_stat(area)) + 
  geom_circlepack_text(vjust = 1.2, color = "grey65",
                       aes(label = paste(after_stat(percent),
                                         "percent"), 
                           size = stage(after_stat = area,
                                        after_scale = size*.85))) +
  ggstamp::theme_void_fill("whitesmoke") +
  theme(legend.position = "none")
  
pride_index <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-06-11/pride_index.csv')

comm_types <- c("large urban city",
                "medium city",
                "small city",
                "small town",
                "very small town",
                "rural community")

pride_it <- pride_index %>% 
  # fix a typo
  mutate(campus_location = str_replace(campus_location, "Swarrthmore", "Swarthmore")) %>% 
  mutate(community_type = fct_relevel(community_type, comm_types)) %>% 
  mutate(state = str_sub(campus_location, -2, -1)) %>% 
  mutate(campus_name = str_replace(campus_name, "University", "U"))

ggplot(pride_it) +
  aes(id = campus_name) + 
  aes(area = students) + 
  aes(fill = rating == 5) +
  geom_circlepack(linewidth = 0.2, color = "grey99") +
  aes(label = str_wrap(after_stat(id), 10)) +
  stat_circlepack_center(geom = GeomText, 
                         lineheight = .8) +
  scale_size(range = c(0,2)) +
  aes(size = students*as.numeric(rating == 5)) + # freshly working
  facet_wrap(facets = vars(community_type))
  
```

```{r}
gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  geom_circlepack(alpha = .5) + 
  coord_equal() +
  aes(area = pop) + 
  geom_circlepack_text(aes(label = after_stat(
    paste(id, "\n",
    round(area/1000000, 1), "mil."))), lineheight = .8)

gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = continent) + 
  geom_circlepack() +
  geom_circlepack_text(alpha = .5) + 
  coord_equal() +
  aes(fill = continent)

last_plot() + 
  aes(id = country)

last_plot() + 
  aes(area = pop)
 

last_plot() + 
  facet_wrap(~continent)


gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) + 
  geom_circlepack() +
  geom_circlepack_text(alpha = .5) +
  aes(area = pop) + 
  coord_equal() + 
  aes(fill = continent) + 
  aes(render = pop > 20000000)

prop_by <- function(...) {
  area <- evalq(area, parent.frame())
  if (length(list(...)) == 0) {
    area / sum(area)
  } else {
    area / ave(area, ..., FUN = function(x) sum(abs(x)))
  }
}

last_plot() +
    aes(label = after_stat(scales::percent(prop_by("All"), 2)))

last_plot() +
    aes(label = after_stat(scales::percent(prop_by(fill), 2)))




```


```{r}


# GeomTextRepel
gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) + 
  geom_circlepack() +
  layer(geom = ggrepel::GeomTextRepel, 
        stat = StatCirclepackcenter, 
        position = "identity") +
  aes(area = pop) + 
  coord_equal() + 
  aes(fill = continent) + 
  aes(render = pop > 20000000)


```



```{r}
gapminder::gapminder %>%
  ggplot() + 
  aes(id = country, area = pop) + 
  geom_circlepack(fun = mean) + 
  geom_circlepack_text(fun = mean) + 
  aes(label = paste(after_stat(id), "\n",
                    round(after_stat(area)/ 1000000), "million")) +
  labs(title = "Average Population - Millions from 1952-2012") + 
  aes(fill = I("plum2")) + 
  coord_equal()
```


```{r}
mpg %>% 
  ggplot() + 
  aes(id = cyl, area = hwy) +
  geom_circlepack(fun = mean) + 
  geom_circlepack_text(fun = mean, lineheight = .8) + 
  scale_size(range = c(4,5)) + 
  labs(title = "Average highway MPG by number of cyl") + 
  coord_equal() + 
  aes(label = 
        after_stat(paste0(id,"cyl\n", 
                          round(area), "mpg"))) + 
  aes(fill = after_stat(area)) + 
  scale_fill_viridis_c(begin = .3, end = .7) 
```


# Package the functions

```{r}
knitrExtra:::chunk_to_r(chunk_name = "compute_panel_circlepack")
knitrExtra:::chunk_to_r(chunk_name = "geom_circlepack")
knitrExtra:::chunk_to_r(chunk_name = "compute_panel_circlepack_center")
knitrExtra:::chunk_to_r(chunk_name = "geom_circlepack_text")
```


```{r, eval = F}
devtools::document()
devtools::check()
devtools::install(pkg = ".", upgrade = "never") 
```


# Issues

Wish list for ggcirclepack: 

## More computation under the hood for a count data case.  

```{r, eval = F}
tidytitanic::tidy_titanic %>% 
  head()
```

```{r}
tidytitanic::tidy_titanic %>% 
  ggplot() + 
  aes(id = "all") + 
  geom_circlepack() +
  geom_circlepack_text(aes(label = after_stat(area)), color = "gray50") +
  coord_equal() + 
  labs(title = "Titanic Passengers")

layer_data(i = 2)

last_plot() +
  aes(fill = sex) +
  scale_size(range = c(3, 6))

last_plot() + 
  aes(alpha = survived) +
  scale_alpha_discrete(range = c(.6,.9))

last_plot() + 
  facet_wrap(~class)

last_plot() + 
  facet_grid(age ~ class)

```

