---
output: 
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggcirclepack

<!-- badges: start -->
<!-- badges: end -->

circle pack is an experimental package that uses the {packcircles} package to handle circle packing computation. 

### Note to the reader

Your feedback is on this work is greatly appreciated.

Beyond the descriptions of our work, we interject comments on our hesitations 🤔 and areas that need some work 🚧, for your consideration marked with emoji.

Your help and feedback would be greatly appreciated on any of the questions...

- Are functions named intuitively? *'According to IBM studies, intuitive variable naming contributes more to code readability than comments, or for that matter, any other factor' McConnell, S. Code complete*
- Do functions work as you expect? 
- Is there rewriting that could make the code more concise?
- What tests should be performed?


# status quo w/o {ggcirclepack}: precomputation required to two more data frames

```{r}
library(tidyverse)
gapminder::gapminder %>%  
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  select(country, pop) %>% 
  mutate(id = row_number()) ->  
df_w_id

packcircles::circleProgressiveLayout(df_w_id$pop,  
                                         sizetype = 'area') ->  
x0y0radius  

x0y0radius %>%  
  packcircles::circleLayoutVertices(npoints = 50) ->  
circle_outlines  

circle_outlines %>% 
  left_join(df_w_id) %>% 
  ggplot() +  
  aes(x = x, y = y) +  
  geom_polygon(colour = "black", alpha = 0.6) +  
  aes(group = id) +  
  aes(fill = pop) +  
  geom_text(data = cbind(df_w_id, x0y0radius),  
            aes(x, y, size = pop, label = country,  
                group = NULL, fill = NULL)) +  
  theme(legend.position = "none") +  
  coord_equal()
```


# Proposed UI

```{r example, out.width="33%", fig.show="hold", message=F, eval = F}
library(tidyverse)
library(ggcirclepack)

gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country, area = pop) +
  geom_circlepack() +               # draws packed circles 
  geom_circlepack_text() +          # labels at the center
  coord_fixed(ratio = 1)
```




# Package functions

## geom_circlepack_text (center)

### Step 1. compute panel

```{r compute_panel_circlepack_center}
#' compute_panel_circlepack_center
#'
#' @return
#' @export
#'
#' @examples
#' # TBD
compute_panel_circlepack_center <- function(data, scales, count = F){

  # data ->
  #   data1
  
  
  
  if(is.null(data$area)){data$area <- 1}
  
   # if(!("weight" %in% names(data))){data$weight <- 1}
  # order matters... Need to add text aesthetics
  if("id" %in% names(data)){data <- group_by(data, id, .add = T)}
  if("fill" %in% names(data)){data <- group_by(data, fill, .add = T)}
  if("alpha" %in% names(data)){data <- group_by(data, alpha, .add = T)}
  if("colour" %in% names(data)){data <- group_by(data, colour, .add = T)}
  if("group" %in% names(data)){data <- group_by(data, group, .add = T)}
  if("linetype" %in% names(data)){data <- group_by(data, linetype, .add = T)}
  if("linewidth" %in% names(data)){data <- group_by(data, linewidth, .add = T)}
  
  
  if(count){
  data %>% 
    count(wt = area) %>% 
    rename(area = n) ->
    data
  }

  data %>%
    pull(area) %>%
    packcircles::circleProgressiveLayout(
      sizetype = 'area') %>%
    cbind(data) %>%
    mutate(label = id)

}
```

### Step 1.1 test compute

```{r}
gapminder::gapminder %>%
filter(continent == "Americas") %>%
  filter(year == 2002) %>%
  # input must have required aesthetic inputs as columns
  select(area = pop, id = country) %>%
  compute_panel_circlepack_center() %>%
  head()

gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  select(id = continent) %>% 
  compute_panel_circlepack_center(count = T)

```

### Step 2 and 3 ggproto and geom

```{r geom_circlepack_text}
StatCirclepackcenter <- ggplot2::ggproto(`_class` = "StatCirclepackcenter",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("id"),
                                  compute_panel = compute_panel_circlepack_center,
                                  default_aes = ggplot2::aes(group = after_stat(id),
                                                             size = after_stat(area))
                                  )


#' Title
#'
#' @param mapping
#' @param data
#' @param position
#' @param na.rm
#' @param show.legend
#' @param inherit.aes
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
#' # TBD
geom_circlepack_text <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCirclepackcenter, # proto object from Step 2
    geom = ggplot2::GeomText, # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```


### Step 4. test geom

```{r}
gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country, area = pop) +
  geom_circlepack_text(alpha = .5) + 
  coord_equal() + 
  labs(title = "gapminder 2002 countries")
```





## geom_circlepack

### Step 1. compute_panel

```{r compute_panel_circlepack}
# Step 1
#' compute_panel_circlepack
#'
#' @param data
#' @param scales
#'
#' @return
#' @export
#'
#' @examples
#' TBD
compute_panel_circlepack <- function(data, scales, count = F){

  if(is.null(data$area)){data$area <- 1}
  
    if(is.null(data$area)){data$area <- 1}
  
   # if(!("weight" %in% names(data))){data$weight <- 1}
  # order matters... Need to add text aesthetics
  if("id" %in% names(data)){data <- group_by(data, id, .add = T)}

  if("fill" %in% names(data)){data <- group_by(data, fill, .add = T)}
  if("alpha" %in% names(data)){data <- group_by(data, alpha, .add = T)}
  if("colour" %in% names(data)){data <- group_by(data, colour, .add = T)}
  if("group" %in% names(data)){data <- group_by(data, group, .add = T)}
  if("linetype" %in% names(data)){data <- group_by(data, linetype, .add = T)}
  if("linewidth" %in% names(data)){data <- group_by(data, linewidth, .add = T)}
  
  if(count){
  data %>% 
    mutate(id = as.integer(as.factor(id))) %>% 
    count(wt = area) %>% 
    rename(area = n) ->
    data
  } else { data$id = 1:nrow(data)}
  
  # data %>%
  #   mutate(id = row_number()) ->
  #   data1

  # if(is.null(data$area)){
  # 
  #   data1 %>%
  #     mutate(area = 1) ->
  #     data1
  # 
  # }

  data %>%
    pull(area) %>%
    packcircles::circleProgressiveLayout(
      sizetype = 'area') %>%
    packcircles::circleLayoutVertices(npoints = 50) %>%
    left_join(data) #%>%

}
```

### Step 1.1. test compute

```{r}
gapminder::gapminder %>%
filter(continent == "Americas") %>%
  filter(year == 2002) %>%
  # input must have required aesthetic inputs as columns
  rename(id = country, area = pop) %>%
  compute_panel_circlepack() %>%
  head()

```

### Step 2 & 3 ggproto and geom

```{r geom_circlepack}
StatCirclepack <- ggplot2::ggproto(`_class` = "StatCirclepack",
                                  `_inherit` = ggplot2::Stat,
                                  required_aes = c("id"),
                                  compute_panel = compute_panel_circlepack,
                                  default_aes = ggplot2::aes(group = after_stat(id))
                                  )

#' Title
#'
#' @param mapping
#' @param data
#' @param position
#' @param na.rm
#' @param show.legend
#' @param inherit.aes
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
#' # TBD
geom_circlepack <- function(mapping = NULL, data = NULL,
                           position = "identity", na.rm = FALSE,
                           show.legend = NA,
                           inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatCirclepack, # proto object from Step 2
    geom = ggplot2::GeomPolygon, # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

### Step 4. test geom

```{r}


gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  geom_circlepack(alpha = .5) + 
  coord_equal() + 
  labs(title = "gapminder 2002 countries")

last_plot() +
  aes(fill = continent) + 
  labs(title = "from 5 continents")

last_plot() +
  aes(area = pop) + 
  geom_circlepack_text() + 
  labs(title = "with very different populations")

last_plot() +
  facet_wrap(facets = vars(continent)) + 
  labs(title = "faceting")

last_plot() + 
  scale_size_continuous(range = c(0, 4)) + 
  theme(legend.position = "none") + 
  labs(title = "remove legends")

last_plot() + 
  aes(area = gdpPercap*pop) + 
  labs(title = "and very different GDPs")

last_plot() + 
  aes(area = gdpPercap) + 
  labs(title = "and per capita GDPs")


gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  geom_circlepack(alpha = .5) + 
  coord_equal() +
  aes(area = pop) + 
  geom_circlepack_text(aes(label = after_stat(
    paste(id, "\n",
    round(area/1000000, 1), "mil."))), lineheight = .8)


gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = continent) + 
  geom_circlepack(count = T) +
  geom_circlepack_text(alpha = .5, count = T) + 
  coord_equal() +
  aes(fill = continent)

last_plot() + 
  aes(id = country)

last_plot() + 
  aes(area = pop)
 

last_plot() + 
  facet_wrap(~continent)


layer_data(i = 2)
```


# Package the functions

```{r}
knitrExtra:::chunk_to_r(chunk_name = "compute_panel_circlepack")
knitrExtra:::chunk_to_r(chunk_name = "geom_circlepack")
knitrExtra:::chunk_to_r(chunk_name = "compute_panel_circlepack_center")
knitrExtra:::chunk_to_r(chunk_name = "geom_circlepack_text")
```


```{r, eval = F}
devtools::check()
devtools::install(pkg = ".", upgrade = "never") 
```


# Issues

Wish list for ggcirclepack: 

## More computation under the hood for a count data case.  

```{r, eval = F}
tidytitanic::tidy_titanic %>% 
  head()
```

```{}
tidytitanic::tidy_titanic() + 
  ggplot() + 
  aes(id = "all") + 
  geom_circlepack(alpha = .5) + 
  geom_circlepack_text(aes(label = afterstat(count))) + # automatically labels with count
  aes(linetype = sex) + 
  aes(color = age) + 
  aes(alpha = survived) + 
  facet_wrap(~class)
```

## Quiet the joins.
## create a ggcirclepack()/defaults_circlepack() function for preferred defaults.


```{r}


```

