---
title: "ggcirclepack and beyond"
author: "Dr. Evangeline Reynolds"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      seal: false
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
```

# ggcirclepack and beyond - highly experimental

<!-- badges: start -->
<!-- badges: end -->

---

Critical moment in my extension journey started with TLP "extend your ability to extend".  
--

Message 1: you can be an extender 

--

Message 2: "where" aesthetics need not only be x and y.

--

aes(x0 = ?, y0 = ?)

--

And what ggplot2 actually looks at is computed for you (x and y)

---


Today I'll present on ggcirclepack. 

---

Background: ggplot2 grammar guide walk through (reference from data to viz gallery)

https://evamaerey.github.io/ggplot2_grammar_guide/geoms_single_series.html#77

---

`r flipbookr::chunk_reveal("background")`

```{r background, include=F}
gapminder::gapminder %>%  
  filter(continent == "Americas") %>%  
  filter(year == 2002) %>%  
  select(country, pop) ->  
prep  

packcircles::circleProgressiveLayout(prep$pop,  
                                         sizetype = 'area') ->  
pack  

pack %>%  
  packcircles::circleLayoutVertices(npoints = 50) ->  
circle_outlines  

circle_outlines %>%  
  ggplot() +  
  aes(x = x, y = y) +  
  geom_polygon(colour = "black", alpha = 0.6) +  
  aes(group = id) +  
  aes(fill = factor(id)) +  
  geom_text(data = cbind(prep, pack),  
            aes(x, y, size = pop, label = country,  
                group = NULL, fill = NULL)) +  
  theme(legend.position = "none") +  
  coord_equal()
```


---

This didn't feel like powerful ggplot2 experience

--

packcircles + ggplot2

--

What would ggcirclepack look like? 

---


Here, the id of the circle isn't defined by the center, x0y0, but by its id and area (an algorithm used under the hood to find x0y0 and x and y for perimeter)

--

Working with compute_panel(), ggols


---


## Installation

You can install the development version of ggcirclepack from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("EvaMaeRey/ggcirclepack")
```

## Example


`r flipbookr::chunk_reveal("example")`

```{r example, include = F}
library(ggcirclepack)
library(tidyverse)
library(magrittr)
library(dplyr)

gapminder::gapminder %>%
filter(year == 2002) %>%
  ggplot() +
  aes(id = country) +
  labs(title = "142 gapminder countries in 2002") +
  geom_polygon_circlepack(alpha = .5) + 
  coord_equal() +
  aes(fill = continent) + 
  aes(area = pop) + 
  geom_text_circlepack() +
  scale_size_continuous(range = c(0, 4)) + 
  facet_wrap(facets = vars(continent)) +
  theme(legend.position = "none") + 
  aes(area = gdpPercap*pop) + 
  aes(area = gdpPercap)
```



---
### Implementation: compute_panel_circle_pack

```{r}
readLines("../R/geom_circle_pack.R")[30:51]
```


---
### Implementation: ggproto

```{r}
readLines("../R/geom_circle_pack.R")[55:62]
```

---
### Implementation: geom_polygon_circlepack

```{r}
readLines("../R/geom_circle_pack.R")[97:110]
```


---

Wish list for ggcirclepack:

More computation under the hood for a count data case.  

```{r}
tidytitanic::tidy_titanic %>% 
  head()
```



---

`r flipbookr::chunk_reveal("titanic", display = "code")`

```{r titanic, eval = F, echo = F}
tidytitanic::tidy_titanic() + 
  ggplot() + 
  aes(x = "all") + 
  geom_polygon_circlepack_count(alpha = .5) + 
  geom_text_circlepack_count() + # automatically labels with count
  aes(linetype = sex) + 
  aes(color = age) + 
  aes(alpha = survived) + 
  facet_wrap(~class)
```

---

Other wish list item


--

Include other circlepacking algorithms?

--

get feedback from {packcircles} author!  

---

# 'Circles are just the best'  TLP


--

# If you squint, a Texas and a Circle look like pretty much the same thing.




---

### Extremely highly experimental: maps and atlasses.




--

aes(where = ?)
--
aes(country = ?)

aes(state = ?)

aes(fips = ?) # us county codes

aes(brain_segment = ?)

aes(tissue = ?)


---

# U.S. States 

- inheriting from geom_polygon
- 'computing' state xy perimeter based aes = my_state_var
- Ryan Miller tutorial start point.


https://evamaerey.github.io/mytidytuesday/2022-11-04-brain/brain.html

---


https://evamaerey.github.io/mytidytuesday/2023-03-06-us-states/us_states.html




---

# U.S. counties

- inheriting from geom_sf()
- 'computing' fips sf geometries column (xy perimeters) based on aes = my_fips_code
- Wilke example start point


https://evamaerey.github.io/mytidytuesday/2023-03-10-ggfips/ggfips_w_sf.html




---

# Brain seg

- inheriting from geom_sf() 
- 'computing' tissue space based on aes = my_seg_name variable
- atlases from ggseg


https://evamaerey.github.io/mytidytuesday/2023-03-12-ggbrain-seg-sf/ggbrain_seg_sf.html





---

```{r, }
library(gganatogram)

a <- gganatogram(data=hgFemale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour") 

a + 
  ggstamp::stamp_wash() +
  geom_path(data = hgFemale_list$brain, aes(x = x, y = -y), color = "red", size = .5, linetype = "dashed") +
  coord_equal()


hgFemale_list$liver %>% 
  rename(organ = id) %>% 
  left_join(hgFemale_key) ->
liver_with_other_info
```


